---
title: "Example Project"
author: "Steven Smith"
date: "4/6/2021"
output: html_document
---

### Analysis of mushroom growth rates

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(broom)
library(modelr)
library(GGally)
dfmush <- read_csv("../../Data_Course_SMITH/Data/mushroom_growth.csv")
```

This data set was collected to determine the fastest conditions for growing Oyster Mushrooms for a farmer's market.

The variables measured were:

```{r,echo=FALSE,message=FALSE,warning=FALSE}
names(dfmush)
```

Take a quick look at our data

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=4,fig.width=4}
ggpairs(dfmush)
```

Now, the analysis! First we want to fit a couple different models:

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod1 <- glm(data = dfmush,
            formula = GrowthRate ~ Light * Humidity)

mod2 <- glm(data = dfmush,
            formula = GrowthRate ~ Light * Nitrogen * Species)

```

Our model formulae are:

```{r,echo=FALSE,message=FALSE}
paste0("MOD1: ", mod1$call[2])
paste0("MOD2: ", mod2$call[2])
```

Which one is better?

```{r,echo=FALSE,message=FALSE}
rmse_mod1 <- residuals(mod1)^2 %>% mean() %>% sqrt()
rmse_mod2 <- residuals(mod2)^2 %>% mean() %>% sqrt()

paste0("The root MSE for MOD1: ",round(rmse_mod1,3))
paste0("The root MSE for MOD2: ",round(rmse_mod2,3))
```

Our first model appears to do a better and simpler job of explaining growth rate, so we will stick with that one

```{r,echo=FALSE,fig.width=4,fig.height=4}
add_predictions(dfmush, mod1) %>% 
  ggplot(aes(x=Light,color=Humidity)) +
  geom_point(aes(y=GrowthRate),alpha=.25) +
  geom_point(aes(y=pred),size=5) +
  theme_bw() +
  labs(caption = "Large points aare model predictions")
```

It appears that for each species, increasing light amount will yield faster growth

```{r,echo=FALSE,message=FALSE}
kableExtra::kable(tidy(mod1))

```

