---
title: "Project3100"
author: "Steven Smith"
date: "4/22/2021"
output: html_document
---

<style type="text/css">
.main-container {
  max-width:1800px;
  margin-left: auto;
  margin-right; auto;
}
</style>

<style>
div.green { background-color:#32ab34; border-radius: Spx; padding: 20px;}
</style>
<div class = "green">

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(tidyverse)
library(patchwork)
library(tidyr)
library(ggplot2)
library(skimr)
library(GGally)
library(dplyr)
library(broom)
library(MASS)
library(plotly)
library(gplots)
library(fitdistrplus)
library(modelr)
library(report)

d1 <- read_xlsx("./FinalProj_Data_1.xlsx")
d1 <- d1 %>% filter(Season != "Career")
d2 <- read_xlsx("./FinalProj_Data_2.xlsx")
d2 <- d2 %>% filter(Season != "Career")
d3 <- read_xlsx("./FinalProj_Data_3.xlsx")
d3 <- d3 %>% filter(Season != "Career")
d4 <- read_xlsx("./FinalProj_Data_4.xlsx")
d4 <- d4 %>% filter(Season != "Career")
d5 <- read_xlsx("./FinalProj_Data_5.xlsx")
d5 <- d5 %>% filter(Season != "Career")
d6 <- read_xlsx("./FinalProj_Data_6.xlsx")
d6 <- d6 %>% filter(Season != "Career")
d7 <- read_xlsx("./FinalProj_Data_7.xlsx")
d7 <- d7 %>% filter(Season != "Career")
d8 <- read_xlsx("./FinalProj_Data_8.xlsx")
d8 <- d8 %>% filter(Season != "Career")
d9 <- read_xlsx("./FinalProj_Data_9.xlsx")
d9 <- d9 %>% filter(Season != "Career")
d10 <- read_xlsx("./FinalProj_Data_10.xlsx")
d10 <- d10 %>% filter(Season != "Career")
d11 <- read_xlsx("./FinalProj_Data_11.xlsx")
d11 <- d11 %>% filter(Season != "Career")
d12 <- read_xlsx("./FinalProj_Data_12.xlsx")
d12 <- d12 %>% filter(Season != "Career")
d13 <- read_xlsx("./FinalProj_Data_13.xlsx")
d13 <- d13 %>% filter(Season != "Career")
d14 <- read_xlsx("./FinalProj_Data_14.xlsx")
d14 <- d14 %>% filter(Season != "Career")
d15 <- read_xlsx("./FinalProj_Data_15.xlsx")
d15 <- d15 %>% filter(Season != "Career")

d1$Player <- "Tobias Harris"
d2$Player <- "Bradley Beal"
d3$Player <- "Jonas Valanciunas"
d4$Player <- "Ish Smith"
d5$Player <- "Danilo Gallinari"
d6$Player <- "Damian Lillard"
d7$Player <- "Nikola Vucevic"
d8$Player <- "Rajon Rondo"
d9$Player <- "Chris Paul"
d10$Player <- "Andre Drummond"
d11$Player <- "Jeff Teague"
d12$Player <- "DeMar DeRozan"
d13$Player <- "Ricky Rubio"
d14$Player <- "Gordon Hayward"
d15$Player <- "Kemba Walker"

Full_Data <- rbind(d1,d2,d3,d4,d5,d6,d7,d8,d9,d10,d11,d12,d13,d14,d15)

cols_to_change_numeric <- c("G","GS","MP","FG","FGA","FG%","3P","3PA","3P%","2P","2PA","2P%","eFG%","FT","FTA","FT%","ORB","DRB","TRB","AST","STL","BLK","TOV","PF","PTS")

Full_Data <- Full_Data %>%
  mutate(Season = str_split(Season,"-") %>% map_chr(1) %>% as.numeric())

Full_Data <- Full_Data %>% 
  as_tibble() %>% 
  mutate(across(cols_to_change_numeric,as.numeric))

names(Full_Data) <- names(Full_Data) %>% make_clean_names()




```
### A Look at Player Shooting per Season
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=fga,color="Red"),alpha=.5) +
  geom_point(aes(y=fg),color="Green",size=1.5) + theme_minimal() + geom_smooth(aes(y=fg),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) + facet_wrap(~player) + geom_smooth(aes(y=fga),method="lm")+
  labs(y= "Field Goals Made and Attempted per Season")
```

### A look into overall shooting
```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod_fg_percent <- glm(data=Full_Data,
            formula = fg_percent ~ season * player * fga)
sqrt(mean(residuals(mod_fg_percent)^2))
summary(mod_fg_percent)
```
### 3-Point Shooting for all players over their seasons
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  geom_point(aes(y=x3pa),color="Red") + geom_point(aes(y=x3p),color="Green") +
  theme_minimal() +  geom_smooth(aes(y=x3p_percent),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "3-Pointers Taken, Made, and %")
```
### Now for each player
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  geom_point(aes(y=x3pa),color="Red") + geom_point(aes(y=x3p),color="Green") +
  theme_minimal() + facet_wrap(~player,scales = "free_y") +  geom_smooth(aes(y=x3p_percent),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "3-Pointers Taken, Made, and %")
```

### Shooting Percentages for players
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  theme_minimal() + facet_wrap(~player) +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "Three Point %")
```
___


## Analyzing interaction of season and player on 3-point percentages. Significance and residual
```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod1 <- glm(data=Full_Data,
            formula = x3p_percent ~ season * player)
summary(mod1)
sqrt(mean(residuals(mod1)^2))
```
### "add_predictions
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
add_predictions(Full_Data,mod1) %>% 
  ggplot(aes(x=season)) + geom_point(aes(y=x3p_percent),alpha=.5) +
  geom_point(aes(y=x3p_percent),color="Green",size=1.5) +
  geom_point(aes(y=pred),color="Red",size=2,alpha=.5) +
  facet_wrap(~player) +
  labs(y= "Three Point %")
```

#### Analyzing interaction of season, player, and 3-pointers attempted on 3-point percentages. Significance and residual
```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod2 <- glm(data=Full_Data,
            formula = x3p_percent ~ season * player * x3pa)
sqrt(mean(residuals(mod2)^2))
summary(mod2)
```

#### graph
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(mod2,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  theme_minimal() + facet_wrap(~player) +  geom_smooth(aes(y=x3p_percent),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "Three Point %")
```

##### 3-point percentages mod2
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(mod2,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  geom_point(aes(y=x3pa),color="Red") +
  theme_minimal() + facet_wrap(~player,scales = "free_y") +  geom_smooth(aes(y=x3p_percent),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "3-Pointers Taken and Made")
```

###add_predictions
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
add_predictions(Full_Data,mod2) %>% 
  ggplot(aes(x=season)) + geom_point(aes(y=x3p_percent),alpha=.5) +
  geom_point(aes(y=x3p_percent),color="Green",size=1.5) +
  geom_point(aes(y=pred),color="Red",size=2,alpha=.5) +
  facet_wrap(~player) +
  labs(y= "Three Point %")
```

```{r,echo=FALSE,message=FALSE,warning=FALSE}
df1 <- Full_Data %>% 
  add_predictions(mod2) 
df1[,c("x3p_percent","pred")] %>% head()
players <- rep(unique(Full_Data$player),each=4)
length(players)
seasons <- rep(c(2021,2022,2022,2023),15)
x3pa <- Full_Data

threepointattempt <- Full_Data %>% 
  dplyr::group_by(player) %>% 
  summarize(mean3pa = mean(x3pa,na.rm = TRUE))
threepa <- rep(threepointattempt$mean3pa,each=4)


newdf = data.frame(season = c(2021,2022,2022,2023),
                   player = players,
                   x3pa = threepa)
```
### Predictions
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=6,fig.width=6}
add_predictions(newdf,mod2) %>% 
  ggplot(aes(x=season,y=pred)) +
  geom_point() + geom_smooth(method="lm") + theme(axis.text.x=element_text(angle=60,hjust=1)) +
  facet_wrap(~player,scales = "free")
```
### Interactions of player and minutes per game + season on 3-point %
```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod3 <- glm(data=Full_Data,
            formula = x3p_percent ~ player * mp + season)
summary(mod3)
sqrt(mean(residuals(mod3)^2))
```
### Minutes vs 3-point %
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=6,fig.width=6}
ggplot(mod3,aes(x=season)) + geom_point(aes(y=mp),color="Blue",alpha=.25) +
  theme_minimal() + facet_wrap(~player) +  geom_smooth(aes(y=mp),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "Minutes per Game")

ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  theme_minimal() + facet_wrap(~player) +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "Three Point %")
```
