---
title: "Project3100"
author: "Steven Smith"
date: "4/27/2021"
output: html_document
---

<style type="text/css">
.main-container {
  max-width:1800px;
  margin-left: auto;
  margin-right; auto;
}
</style>

<style>
div.green { background-color:#32ab34; border-radius: Spx; padding: 20px;}
</style>
<div class = "green">

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(tidyverse)
library(patchwork)
library(tidyr)
library(ggplot2)
library(skimr)
library(GGally)
library(dplyr)
library(broom)
library(MASS)
library(plotly)
library(gplots)
library(fitdistrplus)
library(modelr)
library(report)

d1 <- read_xlsx("./FinalProj_Data_1.xlsx")
d1 <- d1 %>% filter(Season != "Career")
d2 <- read_xlsx("./FinalProj_Data_2.xlsx")
d2 <- d2 %>% filter(Season != "Career")
d3 <- read_xlsx("./FinalProj_Data_3.xlsx")
d3 <- d3 %>% filter(Season != "Career")
d4 <- read_xlsx("./FinalProj_Data_4.xlsx")
d4 <- d4 %>% filter(Season != "Career")
d5 <- read_xlsx("./FinalProj_Data_5.xlsx")
d5 <- d5 %>% filter(Season != "Career")
d6 <- read_xlsx("./FinalProj_Data_6.xlsx")
d6 <- d6 %>% filter(Season != "Career")
d7 <- read_xlsx("./FinalProj_Data_7.xlsx")
d7 <- d7 %>% filter(Season != "Career")
d8 <- read_xlsx("./FinalProj_Data_8.xlsx")
d8 <- d8 %>% filter(Season != "Career")
d9 <- read_xlsx("./FinalProj_Data_9.xlsx")
d9 <- d9 %>% filter(Season != "Career")
d10 <- read_xlsx("./FinalProj_Data_10.xlsx")
d10 <- d10 %>% filter(Season != "Career")
d11 <- read_xlsx("./FinalProj_Data_11.xlsx")
d11 <- d11 %>% filter(Season != "Career")
d12 <- read_xlsx("./FinalProj_Data_12.xlsx")
d12 <- d12 %>% filter(Season != "Career")
d13 <- read_xlsx("./FinalProj_Data_13.xlsx")
d13 <- d13 %>% filter(Season != "Career")
d14 <- read_xlsx("./FinalProj_Data_14.xlsx")
d14 <- d14 %>% filter(Season != "Career")
d15 <- read_xlsx("./FinalProj_Data_15.xlsx")
d15 <- d15 %>% filter(Season != "Career")

d1$Player <- "Tobias Harris"
d2$Player <- "Bradley Beal"
d3$Player <- "Jonas Valanciunas"
d4$Player <- "Ish Smith"
d5$Player <- "Danilo Gallinari"
d6$Player <- "Damian Lillard"
d7$Player <- "Nikola Vucevic"
d8$Player <- "Rajon Rondo"
d9$Player <- "Chris Paul"
d10$Player <- "Andre Drummond"
d11$Player <- "Jeff Teague"
d12$Player <- "DeMar DeRozan"
d13$Player <- "Ricky Rubio"
d14$Player <- "Gordon Hayward"
d15$Player <- "Kemba Walker"

Full_Data <- rbind(d1,d2,d3,d4,d5,d6,d7,d8,d9,d10,d11,d12,d13,d14,d15)

cols_to_change_numeric <- c("G","GS","MP","FG","FGA","FG%","3P","3PA","3P%","2P","2PA","2P%","eFG%","FT","FTA","FT%","ORB","DRB","TRB","AST","STL","BLK","TOV","PF","PTS")

Full_Data <- Full_Data %>%
  mutate(Season = str_split(Season,"-") %>% map_chr(1) %>% as.numeric())

Full_Data <- Full_Data %>% 
  as_tibble() %>% 
  mutate(across(cols_to_change_numeric,as.numeric))

names(Full_Data) <- names(Full_Data) %>% make_clean_names()

```
### In this project, I wanted to look and see if 3-point percentages had increased over the years, as it seems volume certainly has. So I randomly selected 15 players who have played at least 8 seasons, have a 15-minute/game average for their careers and average at least  7.0 points per game. Random.org was used to generate the team to pull from, then again to select a player.

### Just a quick look at player shooting per season

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=fga,color="Red"),alpha=.5) +
  geom_point(aes(y=fg),color="Green",size=1.5) + theme_minimal() + geom_smooth(aes(y=fg),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) + facet_wrap(~player) + geom_smooth(aes(y=fga),method="lm") + theme(legend.position="none") +
  labs(y= "Field Goals Made and Attempted per Season", x="Season")
```

##### Red shows field goal attempts and green shows field goals made

### A look into overall shooting using the formula below finds no significance between the interaction of season, player and field goal attempts on percentages

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod_fg_percent <- glm(data=Full_Data,
            formula = fg_percent ~ season * player * fga)
sqrt(mean(residuals(mod_fg_percent)^2))
formula(mod_fg_percent)
```

### 3-Point Shooting for all players over their seasons. We see a shooting volume increase, which help foment the hypothesis that the league might be shooting better from distance

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  geom_point(aes(y=x3pa),color="Red") + geom_point(aes(y=x3p),color="Green") +
  theme_minimal() +  geom_smooth(aes(y=x3p_percent),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "3-Pointers Taken, Made, and %",x="Season")
```

#### Red = 3-pointers attempted, green = 3-pointers made, and the blue line indicates 3-point %. For percentages 1 on the graph would = 100% shooting

### Now for each player

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  geom_point(aes(y=x3pa),color="Red") + geom_point(aes(y=x3p),color="Green") +
  theme_minimal() + facet_wrap(~player,scales = "free_y") +  geom_smooth(aes(y=x3p_percent),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "3-Pointers Taken, Made, and %", x="Season")
```

#### Red = 3-pointers missed, green = 3-pointers made, and the blue line indicates 3-point %

### 3-Point percentages for players
##### A better look at percentages

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  theme_minimal() + facet_wrap(~player) +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "Three Point %",x="Season")
```

##### 0.1 is 10% and 0.5 Is 50% shooting


## Analyzing interaction of season and player on 3-point percentages. Significance and residual

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod1 <- glm(data=Full_Data,
            formula = x3p_percent ~ season * player)
formula(mod1)
sqrt(mean(residuals(mod1)^2))
```

### Predictions of the above formala

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
add_predictions(Full_Data,mod1) %>% 
  ggplot(aes(x=season)) + geom_point(aes(y=x3p_percent),alpha=.5) +
  geom_point(aes(y=x3p_percent),color="Green",size=1.5) +
  geom_point(aes(y=pred),color="Red",size=2,alpha=.5) +
  facet_wrap(~player) +
  labs(y= "Three Point %",x="Season")
```

##### Predictions in red and 3-point percentage is green

___
### Now, analyzing interaction of season, player, and 3-pointers attempted on 3-point percentages. Significance and residual

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod2 <- glm(data=Full_Data,
            formula = x3p_percent ~ season * player * x3pa)
sqrt(mean(residuals(mod2)^2))
summary(mod2)

```

##### This is the mod that displays great significance of our categories' correlation to 3-point shooting and the mod shows good residuals  compared to other categorical interactions. 'Players' shows a lot of significance, as well as certain seasons they played.

### 3-Point % vs season

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(mod2,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  theme_minimal() + facet_wrap(~player) +  geom_smooth(aes(y=x3p_percent),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "Three Point %",x="Season")
```

##### Added trend line. We see mostly either positive or flat slopes, a couple negative. This looks like it supports the idea that 3-point percentages may be increasing through the seasons.

### 3-point percentages and attempts per season

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
ggplot(mod2,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  geom_point(aes(y=x3pa),color="Red") +
  theme_minimal() + facet_wrap(~player,scales = "free_y") +  geom_smooth(aes(y=x3p_percent),method="lm") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "3-Pointers Taken and Percentage", x="Season")
```

##### Red = 3-pointers attempted and blue shows the percentages for each season

##### We definitely see volume increases. But maybe it just became mathematically better to shoot more threes as long as percentages didn't dip, but we do see some more slightly positive increases. Perhaps it was purely mathematical to green light the shooting of more 3-pointers, and players naturally practiced them more and became better at them.

### Predictions using the above interacting formula

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=5,fig.width=5}
p1 <- add_predictions(Full_Data,mod2) %>% 
  ggplot(aes(x=season)) + geom_point(aes(y=x3p_percent),alpha=.5) +
  geom_point(aes(y=x3p_percent),color="Green",size=1.5) +
  geom_point(aes(y=pred),color="Red",size=2,alpha=.5) +
  facet_wrap(~player) +
  labs(y= "Three Point %", x= "Season")

ggplotly(p1)
```

##### Added predictions = red and green shows 3-point % for each season
##### We see some upwards trends for specific players. Drummond's volume is to low to really consider for making accurate predictions

```{r message=FALSE, warning=FALSE, include=FALSE, r,echo=FALSE}
df1 <- Full_Data %>% 
  add_predictions(mod2) 
df1[,c("x3p_percent","pred")] %>% head()
players <- rep(unique(Full_Data$player),each=4)
length(players)
seasons <- rep(c(2021,2022,2023,2024),15)
seasons <- as.numeric(seasons)
x3pa <- Full_Data

threepointattempt <- Full_Data %>% 
  dplyr::group_by(player) %>% 
  summarize(mean3pa = mean(x3pa,na.rm = TRUE))
threepa <- rep(threepointattempt$mean3pa,each=4)


newdf = data.frame(season = c(2021,2022,2023,2024),
                   player = players,
                   x3pa = threepa)
```

### Predicting into the next 4 future seasons

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=6,fig.width=6}
add_predictions(newdf,mod2) %>% 
  ggplot(aes(x=season,y=pred)) +
  geom_point() + geom_smooth(method="lm") + theme(axis.text.x=element_text(angle=60,hjust=1)) +
  facet_wrap(~player,scales = "free") +
  labs(y= "Three Point % Predictions",x="Future Seasons")
```

##### We see some strong looking predictions, but be mindful that these are showing small overall percentage changes, but they give us an obvious trend to spot.

### Interactions of player and minutes per game + season on 3-point %

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod3 <- glm(data=Full_Data,
            formula = x3p_percent ~ player * mp + season)
formula(mod3)
sqrt(mean(residuals(mod3)^2))
```
This mod shows us significance

### Minutes and 3-point % vs Seasons
##### Given a player shoots good or bad, I'd expect a gain or reduction in minutes.

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=4,fig.width=6}
ggplot(mod3,aes(x=season)) + geom_point(aes(y=mp),color="Blue",alpha=.25) +
  theme_minimal() + facet_wrap(~player) +  geom_smooth(aes(y=mp),method="lm") +  
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "Minutes per Game", x="Season")

ggplot(Full_Data,aes(x=season)) + geom_point(aes(y=x3p_percent),color="Blue",alpha=.25) +
  theme_minimal() + facet_wrap(~player) + 
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y= "Three Point %", x="Season",x="Season")
```

##### We see minutes do have an effect 3-point %, but only for certain players. It appears that a consistently decent 3-point % might a good factor into what's keeping the player playing more minutes. A guy like Andre Drummond, is a traditional big and earns minutes by other methods, but limits spearding the floor like a lot of teams like to do these days.
___

### Interactions of player, position, and season on 3-point %
##### Another aspect to consider would be that of position. Primary ball-handlers, typicaly the guard, might see different kind of shots then say a center, who will likely only take open threes. But a center's body type has typically, in the past, not seen them as finesse players with long distant shooting skills.

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod4<- glm(data=Full_Data,
                formula = x3p_percent ~ player * pos * season)
formula(mod4)

sqrt(mean(residuals(mod4)^2))
```

##### What could be done in the future is collect mean data from many players at each position and compare those values since there isn't a big enough pool to draw from within this data-set. 
